---
title: "Reproducible Research Assignment 1"
output: html_document
---

## Loading and preprocessing the data
The dataset, activity.csv was taken from https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip on the 10 June 2015

```{r, echo=TRUE}
setwd("C:/Coursera/RepData_PeerAssessment1")
activity <- read.csv("activity.csv")
```
Pre Process the data - convert dates to usable format
```{r, echo=TRUE}
activity$date <-as.Date(activity$date,format="%Y-%m-%d")
```
Summary of the data
```{r, echo=TRUE}
summary(activity)
```

###The total number of steps taken per day
```{r, echo=TRUE}
TotalDaily <- aggregate(steps ~ date,activity,sum)
TotalDaily
```

###The following histogram shows the distribution of the total number of steps taken per day
```{r, echo=TRUE}
hist(TotalDaily$steps)
```

###The mean number of steps taken each day is:
```{r, echo=TRUE}
mean(TotalDaily$steps)
```

###The median number of steps taken each day is:
```{r, echo=TRUE}
median(TotalDaily$steps)
```

##average daily activity pattern
Create new dataframe with the average number of steps per interval
```{r, echo=TRUE}
DailyAverage <-aggregate(steps ~ interval, activity,mean)
```

###Time Series plot of average number of steps throughout sample
```{r, echo=TRUE}
plot(DailyAverage,type='l')
```

###The 5 minute interval with the highest average number of steps
```{r, echo=TRUE}
DailyAverage[order(DailyAverage$steps,decreasing=T)[1],]
```

##Imputing missing values

###The total number of missing values
```{r,echo=TRUE}
missing<-subset(activity,is.na(activity$steps))
aggregate(missing$steps, by=list(missing$date), FUN=function(x){NROW(x)})
```

###Impute missing values
Impute missing values by substituting missing value with the average from the corresponding intervals of actual values
Create new dataframe: with.imputed which will contain missing values
```{r, echo=TRUE}
with.imputed<-activity
with.imputed$steps[is.na(with.imputed$steps)] <- DailyAverage$steps[is.na(with.imputed$steps)]
```

###Histogram of total daily steps including imputed values
Adding the imputed values increases the number of average days in the middle of the distribution
```{r, echo=TRUE}
TotalDaily.with.imputed <- aggregate(steps ~ date,with.imputed,sum)
hist(TotalDaily.with.imputed$steps)
```

###The mean number of steps taken each day including imputed values is:
```{r, echo=TRUE}
mean(TotalDaily.with.imputed$steps)
```

###The median number of steps taken each day including imputed values is:
```{r, echo=TRUE}
median(TotalDaily.with.imputed$steps)
```
The mean and median have not been significantly influenced by the imputed values

##Activity pattern differences between weekdays and weekends.
###Create new column in dataframe with a factor signifying the day type
```{r, echo=TRUE}
activity$weekday <- as.POSIXlt(activity$date)$wday
activity$wday <- as.factor(ifelse(activity$weekday == 6,"weekend",ifelse(activity$weekday == 0,"weekend", "weekday")))
str(activity)
```

###
```{r, echo=TRUE}
DailyAverage.bywday <-aggregate(steps ~ interval+wday, activity,mean)
library(lattice)
xyplot(steps ~ interval|wday, data=DailyAverage.bywday, type='l', layout=c(1,2)  )
```